### 2025-01-23 skripsi/chapter4.tex
add discussion, add comparison, change conclusion, revise future work

\begin{table}[H]
  \caption{The configuration of experimental scenarios in this study.}\label{table:experimental_scenarios}
  \centering
  \begin{tabularx}{\textwidth}{L{0.75}*{3}{L{1.09}}} % compute these using approx barycentric coords (x + 3y = 4, x = 0.8y)
    \toprule
    Configuration                    & Scenario 1                                          & Scenario 2                                                          & Scenario 3                                 \\
    \midrule
    Data Source                      & Synthetic (computed from equations)                 & Synthetic (computed from equations) with three viscosity values     & ERA5 dataset                               \\\addlinespace[0.5em]
    Data Augmentations               & Perturbations of three different noise levels       & Perturbation of 10\% the mean standard deviation of function values &                                            \\\addlinespace[0.5em]
    Data Sample Size                 & 5000 unique functions for each noise level          & 500 unique functions for each viscosity value                       & 7000 time steps                            \\\addlinespace[0.5em]
    Number of Modes                  &                                                     &                                                                     &                                            \\\addlinespace[0.5em]
    Scaling                          & Standard scaling                                    & Standard scaling                                                    & Standard scaling                           \\\addlinespace[0.5em]
    Dataset Split                    & 80\% Training and 20\% Testing for each noise level & 80\% Training and 20\% Testing for each viscosity value             & 80\% Training and 20\% Testing             \\\addlinespace[0.5em]
    Regularization Parameter (\(C\)) & 1.0                                                 & 1.0                                                                 & 1.0                                        \\\addlinespace[0.5em]
    Kernel                           & RBF (\(\sigma = features \times X.var()\))          & RBF (\(\sigma = features \times X.var()\))                          & RBF (\(\sigma = features \times X.var()\)) \\\addlinespace[0.5em]
    Additional Tests                 & Exact solutions                                     & Exact solutions                                                     &                                            \\\addlinespace[0.5em]
    Interpretation                   & p-matrix and correlation image                      & p-matrix and correlation image                                      & p-matrix and correlation image             \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

The second scenario extends the goal of the first scenario by examining the model and its behavior when learning a nonlinear PDE\@. The data subsets with different viscosity values are intended to allow us to learn more about how this will affect the performance on the exact solutions for lower viscosity numbers. The model will learn to predict future states of systems governed by the Burgers' equation from past states. The model for each viscosity will be used to predict \lccref{eq:burgers_exact_solution} from initial conditions at time equals zero and the corresponding forcing term of a constant value of zero. The exact equation prediction will be done using a rollout method. The initial condition along with the initial forcing term will be given to the model to predict the solution at the first time step. The model will then use the predicted solution with the forcing term at the first time step to predict the solution at the second time step. This is repeated until the stopping condition of time \(t=10\). The forcing term for the exact solution is a constant zero. Along with the rollout of the exact solution, we also perform roll out on one of the test solution.

@@

\subsection{Scenario 1}
\noindent The first scenario results are presented in four parts. The first is the coefficient prediction metrics which is shown in \lccref{table:scenario_1_spectral_metrics}. The table shows the noise level and the performance metrics.
\begin{table}[H]
  \caption{Performance metrics of coefficient prediction in scenario 1 by noise level.}\label{table:scenario_1_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrrrrrr}
    \toprule
    Noise level & MSE   & RMSE & MAE  & R\textsuperscript{2} & sMAPE \\
    \midrule
    5\%         & 0.86  & 0.93 & 0.57 & 0.96                 & 0.31  \\
    10\%        & 2.06  & 1.43 & 0.93 & 0.91                 & 0.46  \\
    50\%        & 15.59 & 3.95 & 3.04 & 0.50                 & 1.04  \\
    \bottomrule
  \end{tabular}
\end{table}

The scenario is implemented as a Jupyter notebook which imports relevant libraries and our SpectralSVR library. The notebook starts with creating the data with the Antiderivative problem subclass. All random processes use the same PyTorch Generator instance with a seed of 42. The data is then augmented by perturbing the function values to create three versions with the noise levels mentioned in the scenario. Looping through each version, the training and testing subsets are randomly sampled to result in 80\% and 20\% of the original dataset respectively. Then the scaling function is fitted on the training subset inputs. And then, both the training and testing inputs are transformed using the fitted scaling function. Next, an instance of SpectralSVR is trained and tested. Finally, the metrics are stored and the next noise level is processed. Next the 5\% noise model is used to predict the antiderivative of a sine function.

The kernel scaling hyperparameter sigma has the same value of 10 for all noise levels. This is because the inputs are all scaled to approximately have a standard deviation of one. The value itself is from the number of features which comes out to 100 total inputs to the LSSVR\@.

\Cref{table:scenario_1_spectral_metrics} shows that the model is able to generalize from the training data and learn the simple linear antiderivative operator. Focusing on the coefficient of determination (R\textsuperscript{2}) scores, we can see that the model is off by 0.04 to the perfect score of 1.0 for the lowest noise level. While increasing noise levels does degrade the performance, this is partly due to the target values also having been perturbed. This fact can be seen if we look at how predictions from the very same trained models on noisy inputs are compared to the noise-free versions of the antiderivative functions which is shown in \lccref{table:scenario_1_clean_spectral_metrics}. The metrics show that the model predictions are slightly closer to the unperturbed function coefficients compared to the perturbed versions. This can be seen as the influence of the independent synthetic measurement noise in the perturbed targets being removed from the testing metrics.

The higher noise levels shows that the model is still relatively accurate. Looking at the R\textsuperscript{2} score for the 50\% noise level, the model is more accurate than a baseline mean value model. The R\textsuperscript{2} score of 0.5 means that the model predictions has roughly half the error the baseline would produce.

@@

For the metrics between the predictions and the noise-free version of the target function values, the results parallel the outcomes of the coefficient predictions. The metrics show that the model performs slightly better with more pronounced differences for the higher noise levels. This echoes our previous statement on the effect of independent \enquote{measurement} noise in the target values on the evaluation metrics.
\begin{table}[H]
  \caption{Performance metrics of function value from evaluated coefficient prediction compared to unperturbed targets in scenario 1 by noise level.}\label{table:scenario_1_clean_function_metrics}
  \centering
  \begin{tabular}{lrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R\textsuperscript{2} & sMAPE \\
    \midrule
    5\%         & 0.03 & 0.18 & 0.14 & 0.97                 & 0.38  \\
    10\%        & 0.07 & 0.27 & 0.21 & 0.93                 & 0.53  \\
    50\%        & 0.37 & 0.61 & 0.49 & 0.62                 & 0.94  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

The predictions of the exact equation is shown in \lccref{fig:antiderivative_exact}. Visually, we can see that as the noise level increases, the model predictions become worse. Another observation is how the higher the noise level, the more the predicted functions become closer to zero. This is explained by the double penalty phenomenon \autocite{lledoScaledependentVerificationPrecipitation2023}. The loss function used for LSSVR penalizes the model for predicting a non-zero value that turns out to be wrong compared to predicting a zero value. This results in a model with predictions that are close to the mean of the training data.

@@

For inverse problems, we simply use the same testing set with the labels now used to predict the features. The model instances are still the same as before. With the implementation shown in \lccref{fig:spectralsvr_inverse_impl}, we run the inverse function for predicting coefficients with 2000 epochs, a learning rate of 0.02, and randomized starting point from a normal distribution multiplied by a factor of 0.05. The results of inverse prediction for all three noise levels are shown in \lccref{table:scenario_1_inv_spectral_metrics}. These metrics were computed after both the target and predicted input functions are scaled back using the same scaling functions they were scaled down with. Because the scales of the input and output functions are different, the absolute metrics are much larger in these results compared to the forward problem results. For reference, the target input coefficients on average range from -695.65 to 694.9. The relative metrics show that for the lower noise levels, the accuracy is still relatively good. However, the values are lower than the values for the forward test. The difference with the antiderivative prediction becomes more pronounced with the higher noise levels. For the highest noise level, the performance has degraded so far that the predictions are worse than a baseline model from the mean of the target input functions. One must also note that despite the much worse results of the other metrics for the highest noise level, at this time we're unable to explain why sMAPE results in roughly the same performance as in the forward problem.

@@

The final observation we make is how for all the p-matrices the pronounced contribution of the input wave number to their corresponding output wave number. This means that the majority of contributions to each output coefficients come from the corresponding input coefficients of the same wave number. Our knowledge on how the simple derivative equation for Fourier series relate the coefficients of the derivative function and the antiderivative function is shown in \lccref{eq:derivative_coeff}. This aligns with the contributions shown by the p-matrices. This confirms that the model is indeed learning the relations that is defined by the derivative equation.

In summary, the model is able to learn the linear relationship defined by the derivative equation. This is true even for high noise levels. The model also learns the actual relationship between each wave number which is shown in the p-matrix. Inverse prediction of the input function is also successfully done for lower noise levels. At high noise levels, the noise overwhelms the model.

\subsection{Scenario 2}
\noindent The second scenario results are presented in two parts which are the coefficient predictions and function value predictions. The functions are first preprocessed in order to format them into current and next values in time as discussed in \lccref{sec:data_transformation}. The sample pairs are first divided into two for the training and testing sets. Each function is then discretized into 200 time steps. Then, the features are put together as the coefficients of the solution and forcing term at the current time step. The labels are the solution at the next time step. Since there is too much data to fit into memory all at once at the end of this process, we randomly sample the feature and label pairs. This result in a maximum of 8000 samples for the training set and 2000 samples for the testing set. As mentioned in \lccref{sec:data_transformation}, the training features are used to fit a scaling function. And then, all features are scaled using the fitted scaling function.

@@

Another observation we can see is that the difference between the target and rollout predictions is much more pronounced for lower viscosity values. This result is aligned with metrics shown in \lccrefs{table:scenario_2_rollout_spectral_metrics,table:scenario_2_rollout_function_metrics}. The absolute metrics show that the models are performing several times worse in rollout compared to the single time step tests shown in \lccrefs{table:scenario_2_spectral_metrics,table:scenario_2_function_metrics}. This seems to be the same challenge that traditional solvers also face with the shocks that may be present with lower viscosity values. This means that for rollout scenarios, the model has a much harder time with lower viscosity values.
\begin{table}[H]
  \caption{Performance metrics of coefficient rollout in scenario 2 by viscosity.}\label{table:scenario_2_rollout_spectral_metrics}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    \(\nu \) & MSE      & RMSE     & MAE      & sMAPE \\
    \midrule
    0.0      & 2.91e-01 & 5.40e-01 & 4.07e-01 & 0.90  \\
    0.01     & 1.61e-01 & 4.01e-01 & 2.86e-01 & 0.75  \\
    0.1      & 1.70e-01 & 4.12e-01 & 3.02e-01 & 0.81  \\
    \bottomrule
  \end{tabular}
\end{table}
\begin{table}[H]
  \caption{Performance metrics of function values evaluated from coefficient rollout in scenario 2 by viscosity.}\label{table:scenario_2_rollout_function_metrics}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    \(\nu \) & MSE      & RMSE     & MAE      & sMAPE \\
    \midrule
    0.0      & 5.81e-02 & 2.41e-01 & 1.89e-01 & 1.04  \\
    0.01     & 3.17e-02 & 1.78e-01 & 1.40e-01 & 0.94  \\
    0.1      & 2.91e-02 & 1.71e-01 & 1.33e-01 & 0.92  \\
    \bottomrule
  \end{tabular}
\end{table}

The exact function \cref{eq:burgers_exact_solution} is computed for each viscosity value. The discrete Fourier transform of the values are then used for doing rollout with the model. For the forcing term, we just use a constant function with a value of zero. The rollout predictions are shown in \lccref{fig:scenario_2_exact}. The model is not successful in the rollout for the inviscid equation and for the viscosity of \(\nu=0.01\). Both of these cases, the model produces too much error that the original function is no longer recognizable in the predictions. For the higher viscosity value of \(\nu=0.1\) the rollout produces a recognizable prediction. The error stays to about half the maximum function value. We do see some more noticeable error in the flatter parts of the function. The error also take a very similar shape across the different viscosity values. Since we essentially used the same solutions for training the model for each viscosity, the similarity in error means that it can be alleviated with more diverse training samples. The error for the flatter functions may also be alleviated by including similar flatter functions during training.
\begin{figure}[H]
  \centering
  \begin{adjustwidth}{-0.1\linewidth}{-0.1\linewidth}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_target_0.0.pgf}
      \end{adjustbox}
      \caption{The target for \(\nu=0.0\)}\label{fig:sc2_exact_target_0.0}
    \end{subfigure}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_pred_0.0.pgf}
      \end{adjustbox}
      \caption{The prediction for \(\nu=0.0\)}\label{fig:sc2_exact_pred_0.0}
    \end{subfigure}
    \begin{subfigure}{0.32\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_diff_0.0.pgf}
      \end{adjustbox}
      \caption{The difference for \(\nu=0.0\)}\label{fig:sc2_exact_diff_0.0}
    \end{subfigure}
    \\[0.7\baselineskip]
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_target_0.01.pgf}
      \end{adjustbox}
      \caption{The target for \(\nu=0.01\)}\label{fig:sc2_exact_target_0.01}
    \end{subfigure}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_pred_0.01.pgf}
      \end{adjustbox}
      \caption{The prediction for \(\nu=0.01\)}\label{fig:sc2_exact_pred_0.01}
    \end{subfigure}
    \begin{subfigure}{0.32\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_diff_0.01.pgf}
      \end{adjustbox}
      \caption{The difference for \(\nu=0.01\)}\label{fig:sc2_exact_diff_0.01}
    \end{subfigure}
    \\[0.7\baselineskip]
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_target_0.1.pgf}
      \end{adjustbox}
      \caption{The target for \(\nu=0.1\)}\label{fig:sc2_exact_target_0.1}
    \end{subfigure}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_pred_0.1.pgf}
      \end{adjustbox}
      \caption{The prediction for \(\nu=0.1\)}\label{fig:sc2_exact_pred_0.1}
    \end{subfigure}
    \begin{subfigure}{0.32\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_diff_0.1.pgf}
      \end{adjustbox}
      \caption{The difference for \(\nu=0.1\)}\label{fig:sc2_exact_diff_0.1}
    \end{subfigure}
    % \\[0.7\baselineskip]
  \end{adjustwidth}
  \caption{The rollout predictions for one of the exact function in \lccref{eq:burgers_exact_solution}. The difference is calculated as the targets subtracted by the predictions.}\label{fig:scenario_2_exact}
\end{figure}

@@

The correlation image and p-matrix for each viscosity can be seen in \lccref{fig:scenario_2_interpretation}. Analyzing the correlation images first, we see that some columns are sorted in correlation to how the samples are sorted based on the real component of the 4th output wave number. For the inviscid equation, this correlation is the strongest with multiple columns exhibiting the correlation. The highest and visually most sorted values are from the forcing term portion of the inputs. Different wave numbers in the inputs are also sorted differently, this indicates an inverse correlation. For a higher viscosity of \(\nu=0.01\), a weaker correlation is still present.

@@

Once we get to the highest viscosity, while the order is still present, we see a more random arrangement of the correlation image values. This observation and the lower scores can both be explained by the level of temporal discretization. This is the same issue we encounter when we use some traditional methods for higher viscosity values with the same time step for lower viscosity values \autocite{kassamFourthOrderTimeSteppingStiff2005,seydaogluNumericalSolutionBurgers2016}. The issue arises in stiff differential equations. Stiff differential equations refer to the presence of a rapidly varying component. We can see that the larger amplitudes of high frequency components in the forcing term to represent the rapidly varying components as shown in \lccrefs{fig:burgers_forcing_0.0,fig:burgers_forcing_0.01,fig:burgers_forcing_0.1}. Traditionally, stiff equations would require specific numerical methods purpose built to efficiently solve the problem. Otherwise, the using a solver like FDM would require much finer time steps than the one we are using here.

The p-matrix shows the difference in contributions of both input functions. For the inviscid equation, the contributions mainly come from the forcing term. The current solution contributes a smaller amount in comparison. There is similarity with the antiderivative p-matrices in this case. The contributions to each output wave number mainly come from input coefficients with the same wave number. However, there is a noticeable contribution from other wave numbers too unlike the antiderivative case. The biggest contribution from the current solution is the constant/bias term. For higher viscosity values, notice that the contribution of the previous solution increases. The increase is more pronounced for higher frequencies. This aligns with the quadratic scaling from the wave number on the coefficients \autocite{canutoSpectralMethodsEvolution2007}. The p-matrix for lower viscosity values show more structured contribution of other wave numbers in comparison to the higher wave numbers. This structure comes in the form of a grid pattern. For the real components of the output, the contribution of other wave numbers oscillate between higher and lower values depending on whether the input component is real or not. The imaginary output components on the other hand have a more constant contribution across the input components. The contribution once again become more pronounced the higher the absolute value of the wave number.

To summarize, Low viscosity values are a challenge to the model as with traditional methods. Despite that, this scenario confirms the model's ability to learn the nonlinear relationship defined by the Burgers' equation. The model is also able to stay stable with a relatively large time step. This concludes the synthetic data portion of the study.

>>>>

\begin{table}[H]
  \caption{The configuration of experimental scenarios in this study.}\label{table:experimental_scenarios}
  \centering
  \begin{tabularx}{\textwidth}{L{0.75}*{3}{L{1.09}}} % compute these using approx barycentric coords (x + 3y = 4, x = 0.8y)
    \toprule
    Configuration                    & Scenario 1                                          & Scenario 2                                                                                            & Scenario 3                                   \\
    \midrule
    Data Source                      & Synthetic (computed from equations)                 & Synthetic (computed from equations) with three viscosity values                                       & ERA5 dataset                                 \\\addlinespace[0.5em]
    Data Augmentations               & Perturbations of three different noise levels       & Perturbation of 10\% the mean standard deviation of function values                                   &                                              \\\addlinespace[0.5em]
    Data Sample Size                 & 5000 unique functions for each noise level          & 500 unique functions for each viscosity value                                                         & 7000 time steps                              \\\addlinespace[0.5em]
    Number of Modes                  &                                                     &                                                                                                       &                                              \\\addlinespace[0.5em]
    Scaling                          & Standard scaling                                    & Standard scaling                                                                                      & Standard scaling                             \\\addlinespace[0.5em]
    Dataset Split                    & 80\% Training and 20\% Testing for each noise level & 80\% Training and 20\% Testing for each viscosity value                                               & 1995{-}2013 Training and 2020{-}2022 Testing \\\addlinespace[0.5em]
    Regularization Parameter (\(C\)) & 1.0                                                 & 1.0                                                                                                   & 1.0                                          \\\addlinespace[0.5em]
    Kernel                           & RBF (\(\sigma = features \times X.var()\))          & RBF (\(\sigma = features \times X.var()\))                                                            & RBF (\(\sigma = features \times X.var()\))   \\\addlinespace[0.5em]
    Additional Tests                 & Exact solutions and Inverse Problem                 & Exact solution in \cref{eq:burgers_exact_solution} and in initial value problem formulation (rollout) &                                              \\\addlinespace[0.5em]
    Comparisons                      &                                                     & Method of Lines with FDM in space                                                                     &                                              \\\addlinespace[0.5em]
    Machine                          & Local CPU                                           & Local CPU                                                                                             & Kaggle CPU                                   \\\addlinespace[0.5em]
    Interpretation                   & p-matrix and correlation image                      & p-matrix and correlation image                                                                        & p-matrix and correlation image               \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

The second scenario extends the goal of the first scenario by examining the model and its behavior when learning a nonlinear PDE\@. The data subsets with different viscosity values are intended to allow us to learn more about how this will affect the performance on the exact solutions for lower viscosity numbers. The model will learn to predict future states of systems governed by the Burgers' equation from past states. The model for each viscosity will be used to predict \lccref{eq:burgers_exact_solution} from initial conditions at time equals zero and the corresponding forcing term of a constant value of zero. The exact equation prediction will be done using a initial value problem formulation, also described as a rollout. The initial solution value along with the forcing term will be given to the model to predict the solution at the first time step. The model will then use the predicted solution with the forcing term to predict the solution at the second time step. This is repeated until the stopping condition of time \(t=10\). The forcing term for the exact solution is a constant zero. In addition to the rollout of the exact solution, we also perform rollout on of the test set. This will be done in comparison to traditional numerical methods, namely the commonly used method of lines in time using SciPy and torchdiffeq and finite differences in space with NumPy gradient and PyTorch gradient respectively. Specifically, the SciPy odeint function uses lsoda from the FORTRAN library odepack. And, the torchdiffeq version will use Implicit Adams-Bashforth-Moulton (IABM) method. Both of these were chosen for their availability for use in Python and therefore the kind of numerical solver one could set up with relative reliability.

@@

\subsection{Scenario 1}
\noindent The scenario is implemented as a Jupyter notebook which imports relevant libraries and our SpectralSVR library. The notebook starts with creating the data with the Antiderivative problem subclass. All random processes use the same PyTorch Generator instance with a seed of 42. The data is then augmented by perturbing the function values to create three versions with the noise levels mentioned in the scenario. Looping through each version, the training and testing subsets are randomly sampled to result in 80\% and 20\% of the original dataset respectively. Then the scaling function is fitted on the training subset inputs. And then, both the training and testing inputs are transformed using the fitted scaling function. Next, an instance of SpectralSVR is trained and tested. Finally, the metrics are stored and the next noise level is processed.

The kernel scaling hyperparameter sigma has the same value of 10 for all noise levels. This is because the inputs are all scaled to approximately have a standard deviation of one. The value itself is from the number of features which comes out to 100 total inputs to the LSSVR\@. This was multiplied with the average standard deviation of all features across samples of 1 and then squared to get our current scaling hyperparameter value of 10.

The first scenario results are presented in four parts. The first is the coefficient prediction metrics which is shown in \lccref{table:scenario_1_spectral_metrics}. The table shows the noise level and the performance metrics. Focusing on the coefficient of determination (R\textsuperscript{2}) scores, we can see that the model is off by 0.04 to the perfect score of 1.0 for the lowest noise level. For higher noise levels, the model performance gets worse with the R\textsuperscript{2} score of 0.5 for the highest noise level.
\begin{table}[H]
  \caption{Performance metrics of coefficient prediction in scenario 1 by noise level.}\label{table:scenario_1_spectral_metrics}
  \centering
  \begin{tabular}{lrrrrrrrrr}
    \toprule
    Noise level & MSE   & RMSE & MAE  & R\textsuperscript{2} & sMAPE \\
    \midrule
    5\%         & 0.86  & 0.93 & 0.57 & 0.96                 & 0.31  \\
    10\%        & 2.06  & 1.43 & 0.93 & 0.91                 & 0.46  \\
    50\%        & 15.59 & 3.95 & 3.04 & 0.50                 & 1.04  \\
    \bottomrule
  \end{tabular}
\end{table}

The metrics against the non perturbed values are shown in \lccref{table:scenario_1_clean_spectral_metrics}. The results show that the model performs best with lower noise levels, with the highest R\textsuperscript{2} value of 0.96 achieved at noise of 5\% average function value standard deviation. Higher noise levels sees the model performing worse with an R\textsuperscript{2} value of 0.62 at the highest noise level.

@@

For the metrics between the predictions and the noise-free version of the target function values shown in \lccref{table:scenario_1_clean_function_metrics}, the results parallel the outcomes of the coefficient predictions. The metrics show that the model performs slightly better with more pronounced differences for the higher noise levels.
\begin{table}[H]
  \caption{Performance metrics of function value from evaluated coefficient prediction compared to unperturbed targets in scenario 1 by noise level.}\label{table:scenario_1_clean_function_metrics}
  \centering
  \begin{tabular}{lrrrrrrr}
    \toprule
    Noise level & MSE  & RMSE & MAE  & R\textsuperscript{2} & sMAPE \\
    \midrule
    5\%         & 0.03 & 0.18 & 0.14 & 0.97                 & 0.38  \\
    10\%        & 0.07 & 0.27 & 0.21 & 0.93                 & 0.53  \\
    50\%        & 0.37 & 0.61 & 0.49 & 0.62                 & 0.94  \\
    \bottomrule
  \end{tabular}
\end{table}

@@

The predictions of the exact equation is shown in \lccref{fig:antiderivative_exact}. Visually, we can see that as the noise level increases, the model predictions become worse. Another observation is how the higher the noise level, the more the predicted functions become closer to zero.

@@

For inverse problems, we simply use the same testing set with the labels now used to predict the features. The model instances are still the same as before. With the implementation shown in \lccref{fig:spectralsvr_inverse_impl}, we run the inverse function for predicting coefficients with 2000 epochs, a learning rate of 0.02, and randomized starting point from a normal distribution multiplied by a factor of 0.05. The results of inverse prediction for all three noise levels are shown in \lccref{table:scenario_1_inv_spectral_metrics}. These metrics were computed after both the target and predicted input functions are scaled back using the same scaling functions they were scaled down with. Because the scales of the input and output functions are different, the absolute metrics are much larger in these results compared to the forward problem results. For reference, the target input coefficients on average range from -695.65 to 694.9. The relative metrics show that for the lower noise levels, the accuracy is still relatively good. However, the values are lower than the values for the forward test. The difference with the antiderivative prediction becomes more pronounced with the higher noise levels. For the highest noise level, the performance has degraded so far that the predictions are worse than a baseline model from the mean of the target input functions.

@@

\subsection{Scenario 2}
\noindent The second scenario results are presented in two parts which are the coefficient predictions and function value predictions. The functions are first preprocessed in order to format them into current and next values in time as discussed in \lccref{sec:data_transformation}. The sample pairs are first divided into two for the training and testing sets. Each function is then discretized into 200 time steps. Then, the features are put together as the coefficients of the solution and forcing term at the current time step. The labels are the solution at the next time step. Since there is too much data to fit into memory all at once at the end of this process, we randomly sample the feature and label pairs. This result in a maximum of 8000 features and labels for the training set and 2000 features and labels from the testing set. As mentioned in \lccref{sec:data_transformation}, the training features are used to fit a scaling function. And then, all features are scaled using the fitted scaling function.

@@

Another observation we can see is that the difference between the target and rollout predictions is much more pronounced for lower viscosity values. This result is aligned with metrics shown in \lccrefs{table:scenario_2_rollout_spectral_metrics,table:scenario_2_rollout_function_metrics}. The absolute metrics show that the models are performing several times worse in rollout compared to the single time step tests shown in \lccrefs{table:scenario_2_spectral_metrics,table:scenario_2_function_metrics}.
\begin{table}[H]
  \caption{Performance metrics of coefficient rollout in scenario 2 by viscosity.}\label{table:scenario_2_rollout_spectral_metrics}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    \(\nu \) & MSE      & RMSE     & MAE      & sMAPE \\
    \midrule
    0.0      & 2.91e-01 & 5.40e-01 & 4.07e-01 & 0.90  \\
    0.01     & 1.61e-01 & 4.01e-01 & 2.86e-01 & 0.75  \\
    0.1      & 1.70e-01 & 4.12e-01 & 3.02e-01 & 0.81  \\
    \bottomrule
  \end{tabular}
\end{table}
\begin{table}[H]
  \caption{Performance metrics of function values evaluated from coefficient rollout in scenario 2 by viscosity.}\label{table:scenario_2_rollout_function_metrics}
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    \(\nu \) & MSE      & RMSE     & MAE      & sMAPE \\
    \midrule
    0.0      & 5.81e-02 & 2.41e-01 & 1.89e-01 & 1.04  \\
    0.01     & 3.17e-02 & 1.78e-01 & 1.40e-01 & 0.94  \\
    0.1      & 2.91e-02 & 1.71e-01 & 1.33e-01 & 0.92  \\
    \bottomrule
  \end{tabular}
\end{table}
The exact function \cref{eq:burgers_exact_solution} is computed for each viscosity value. The discrete Fourier transform of the values are then used for doing rollout with the model. For the forcing term, we just use a constant function with a value of zero. The rollout predictions are shown in \lccref{fig:scenario_2_exact}. The model is not successful in the rollout for the inviscid equation and for the viscosity of \(\nu=0.01\). Both of these cases, the model produces too much error that the original function is no longer recognizable in the predictions. For the higher viscosity value of \(\nu=0.1\) the rollout produces a recognizable prediction. The error stays to about half the maximum function value.
\begin{figure}[H]
  \centering
  \begin{adjustwidth}{-0.1\linewidth}{-0.1\linewidth}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_target_0.0.pgf}
      \end{adjustbox}
      \caption{The target for \(\nu=0.0\)}\label{fig:sc2_exact_target_0.0}
    \end{subfigure}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_pred_0.0.pgf}
      \end{adjustbox}
      \caption{The prediction for \(\nu=0.0\)}\label{fig:sc2_exact_pred_0.0}
    \end{subfigure}
    \begin{subfigure}{0.32\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_diff_0.0.pgf}
      \end{adjustbox}
      \caption{The difference for \(\nu=0.0\)}\label{fig:sc2_exact_diff_0.0}
    \end{subfigure}
    \\[0.7\baselineskip]
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_target_0.01.pgf}
      \end{adjustbox}
      \caption{The target for \(\nu=0.01\)}\label{fig:sc2_exact_target_0.01}
    \end{subfigure}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_pred_0.01.pgf}
      \end{adjustbox}
      \caption{The prediction for \(\nu=0.01\)}\label{fig:sc2_exact_pred_0.01}
    \end{subfigure}
    \begin{subfigure}{0.32\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_diff_0.01.pgf}
      \end{adjustbox}
      \caption{The difference for \(\nu=0.01\)}\label{fig:sc2_exact_diff_0.01}
    \end{subfigure}
    \\[0.7\baselineskip]
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_target_0.1.pgf}
      \end{adjustbox}
      \caption{The target for \(\nu=0.1\)}\label{fig:sc2_exact_target_0.1}
    \end{subfigure}
    \begin{subfigure}{0.33\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_pred_0.1.pgf}
      \end{adjustbox}
      \caption{The prediction for \(\nu=0.1\)}\label{fig:sc2_exact_pred_0.1}
    \end{subfigure}
    \begin{subfigure}{0.32\linewidth}
      \begin{adjustbox}{width=\linewidth}
        \input{figures/burgers_exact_diff_0.1.pgf}
      \end{adjustbox}
      \caption{The difference for \(\nu=0.1\)}\label{fig:sc2_exact_diff_0.1}
    \end{subfigure}
    % \\[0.7\baselineskip]
  \end{adjustwidth}
  \caption{The rollout predictions for one of the exact function in \lccref{eq:burgers_exact_solution}. The difference is calculated as the targets subtracted by the predictions.}\label{fig:scenario_2_exact}
\end{figure}

@@

The correlation image and p-matrix for each viscosity can be seen in \lccref{fig:scenario_2_interpretation}. Analyzing the correlation images first, we see that some columns are sorted in correlation to how the samples are sorted based on the real component of the 4th output wave number. For the inviscid equation, this correlation is the strongest with multiple columns exhibiting the correlation. The highest and visually most sorted values are from the forcing term portion of the inputs. Different wave numbers in the inputs are also sorted differently, this indicates an inverse correlation. For a higher viscosity of \(\nu=0.01\), a weaker correlation is still present. Once we get to the highest viscosity, while the order is still present, we see a more random arrangement of the correlation image values.

@@

The p-matrix shows the difference in contributions of both input functions. For the inviscid equation, the contributions mainly come from the forcing term. The current solution contributes a smaller amount in comparison. There is similarity with the antiderivative p-matrices in this case. The contributions to each output wave number mainly come from input coefficients with the same wave number. However, there is a noticeable contribution from other wave numbers too unlike the antiderivative case. The biggest contribution from the current solution is the constant/bias term. For higher viscosity values, notice that the contribution of the previous solution increases.
